Adapting bootstrap to our problem - Part 2

### 2024-07-11 adapt_boot2.Rmd
Not done yet

---
title: "Adapting bootstrap to our problem"
output: pdf_document
date: "2024-07-10"
---

@@

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
```

@@

```{r}
# Number of bootstrap samples
B <- 500
```

@@

```{r}
# Making of bootstrap samples
df_boot <- f_boot(random_points_1, B)
```

```{r}
# In case B=1
# plot(x = df_boot$ech1.x, y = df_boot$ech1.y)
```

@@

```{r}
f_boot2 <- function(df_sample, nboot){
  # df_sample: original sample
  # nboot: number of bootstrap samples made thanks to df_sample

  odd_even1 <- sample(1:2, 1) # Randomly put more weight on even or odd lines
  odd_even2 <- sample(1:2, 1) # The same for "columns"
  # df_sample$prob <- (as.numeric(floor(df_sample$y/10)%%2 == odd_even1%%2)*(3/4) + as.numeric(floor(df_sample$y/10)%%2 != odd_even1%%2)*(1/4) + 
  #                      df_sample[df_sample$l]$x +
  #                      as.numeric(floor(df_sample$x/10)%%2 == odd_even2%%2)*(3/4) + as.numeric(floor(df_sample$x/10)%%2 != odd_even2%%2)*(1/4))/2

  df_sample$step <- rep(0, nrow(df_sample))
  for (line in unique(df_sample$l)) {
    df_sample$step <- (df_sample$step == 0)*(row(df_sample[df_sample$l == line,])[,1]%%2) + (df_sample$step != 0)*df_sample$step
  }
  # df_sample$prob <- ((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2)
  df_sample$prob <- (((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2) == 0)*(1/4) +
    (((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2) == 1)*(1/2) +
    (((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2) == 2)*(3/4)
  df_sample$l <- NULL
  df_sample$step <- NULL

@@

```{r}
# Making of bootstrap samples
df_boot2 <- f_boot2(random_points, B)
```

```{r}
# In case B=1
# plot(x = df_boot2$ech1.x, y = df_boot2$ech1.y)
```

@@

### First method

@@

```{r}
# Making of one estimation for each sample made by the bootstrap method
est_boot_1 <- est_boot1(df_boot)
```

```{r}
mean(est_boot_1)
# 31055.71 e.g.
# 30488.92
```

```{r}
var(est_boot_1)
# 2791309 e.g.
# 2577277
```

### Second method

```{r}
# Making of one estimation for each sample made by the bootstrap method
est_boot_2 <- est_boot1(df_boot2)
```

```{r}
mean(est_boot_2)
# 31153.14 e.g.
# 30499.57
```

```{r}
var(est_boot_2)
# 3222756 e.g.
# 2351291
```

@@

```{r}
comparison <- function(ncomp, B) {
  # ncomp: number of bootstrap series to compare
  # B: number of bootstrap replications each time
  tot_mean <- c()
  tot_var <- c()
  b <- 1
  for (b in 1:ncomp) {
    ## First method
    # We only keep x and y in the dataframe
    random_points_1 <- random_points
    random_points_1$l <- NULL
    df_boot <- f_boot(random_points_1, B)
    est_boot_1 <- est_boot1(df_boot)

    ## Second method
    df_boot2 <- f_boot2(random_points, B)
    est_boot_2 <- est_boot1(df_boot2)

    tot_mean <- append(tot_mean, as.numeric(abs(mean(est_boot_2)-30942) <  abs(mean(est_boot_1)-30942))*2-1)
    # 1: the second method is better
    tot_var <- append(tot_var, as.numeric(var(est_boot_2) <  var(est_boot_1))*2-1)
    # 1: the second method is better
    b <- b+1
  }
  return(list(tot_mean, tot_var))
}
```

```{r}
res <- comparison(1000, 800)
```

```{r}
# If it is positive, the second method is the best to estimate the total
sum(res[[1]])
# With ncomp = 100 and B = 500
# 10 e.g.
# 18

# With ncomp = 1000 and B = 800
# -4
```

@@

Adapting bootstrap to the problem seems relevant: it does not improve much the estimator but it reduces its variance.

>>>>

---
title: "Adapting bootstrap to our problem - Part 2"
output: pdf_document
date: "2024-07-11"
---

@@

```{r}
library(readxl) # to read the initial file
library(magrittr) # for pipes
library(tibble) # for add_column
library(ggplot2)
```

@@

@@

@@

```{r}
f_boot2 <- function(df_sample, nboot){
  # df_sample: original sample (4 variables: x,y,l,var)
  # nboot: number of bootstrap samples made thanks to df_sample

  odd_even1 <- sample(1:2, 1) # Randomly put more weight on even or odd lines
  odd_even2 <- sample(1:2, 1) # The same for "columns"

  df_sample$step <- rep(0, nrow(df_sample))
  for (line in unique(df_sample$l)) {
    df_sample$step <- as.numeric(df_sample$step == 0)*(row(df_sample[df_sample$l == line,])[,1]%%2) + as.numeric(df_sample$step != 0)*df_sample$step
  }

  df_sample$prob <- (((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2) == 0)*(1/4) +
    (((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2) == 1)*(1/2) +
    (((df_sample$l+odd_even1)%%2 + (df_sample$step+odd_even2)%%2) == 2)*(3/4)
  df_sample$l <- NULL
  df_sample$step <- NULL

@@

@@

@@

@@

```{r}
comparison2 <- function(B) {
  # B: number of bootstrap replications each time
  mean_vec1 <- c()
  mean_vec2 <- c()
  var_vec1 <- c()
  var_vec2 <- c()
  b1 <- 100
  while (b1  < B) {
    ## First method
    # We only keep x and y in the dataframe
    random_points_1 <- random_points
    random_points_1$l <- NULL
    df_boot <- f_boot(random_points_1, b1)
    est_boot_1 <- est_boot1(df_boot)

    ## Second method
    df_boot2 <- f_boot2(random_points, b1)
    est_boot_2 <- est_boot1(df_boot2)

    mean_vec1 <- append(mean_vec1, abs(mean(est_boot_1)-30942))
    mean_vec2 <- append(mean_vec2, abs(mean(est_boot_2)-30942))

    var_vec1 <- append(var_vec1, var(est_boot_1))
    var_vec2 <- append(var_vec2, var(est_boot_2))

    # tot_mean <- append(tot_mean, as.numeric(abs(mean(est_boot_2)-30942) <  abs(mean(est_boot_1)-30942))*2-1)
    # 1: the second method is better
    # tot_var <- append(tot_var, as.numeric(var(est_boot_2) <  var(est_boot_1))*2-1)
    # 1: the second method is better
    b1 <- b1+10
  }
  return(list(mean_vec1, mean_vec2, var_vec1, var_vec2))
}
```

```{r}
res2 <- comparison2(500)
```

```{r}
res <- comparison2(500)
```

```{r}
data = as.data.frame(cbind(1:length(res2[[1]]), res2[[1]], res2[[2]]))
p1 <- ggplot(data, aes(data[,1], data[,2]))
p2 <- ggplot(data, aes(data[,1], data[,3]))
p1 + p2 + geom_point()
p1 + geom_point

# p <- ggplot(data, aes(data[,1], cbind(data[,2], data[,3])))
# p + geom_point()

p <- ggplot(data, aes(data[,1], data[,3] - data[,2]))
p + geom_point()
```

```{r}
data_var = as.data.frame(cbind(1:length(res2[[1]]), res2[[3]], res2[[4]]))

p <- ggplot(data_var, aes(data_var[,1], data_var[,3] - data_var[,2]))
p + geom_point()
```

```{r}
data_var = as.data.frame(cbind(1:length(res2[[1]]), res2[[3]], res2[[4]]))

p <- ggplot(data_var, aes(data_var[,1], data_var[,3] - data_var[,2]))
p + geom_point()
```

```{r}
ggplot(x = 1:length(res[[1]]), y = res[[1]]) +
  plot(x = 1:length(res[[1]]), y = res[[2]], col = "red")
geom_point(
    mapping = aes(y = group_mean), data = group_means_df,
    colour = 'red', size = 3)

```

@@
